---
title: "adhd 8812 1"
format: html
editor: visual
---

ADHD

```{r}
packages <- c(
  "readxl", "dplyr", "tidyr", "stringr",
  "lavaan", "semTools", "psych", "broom"
)

to_install <- packages[!packages %in% rownames(installed.packages())]
if(length(to_install) > 0) install.packages(to_install)

library(readxl)
library(dplyr)
library(tidyr)
library(stringr)
library(lavaan)
library(semTools)
library(psych)
library(broom)
library(dplyr)

data_path <- "/Users/yangsiyutong/Desktop/QuestionPro-SR-RawData-4077617-04-12-2026-T094743.646.xlsx"
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

raw <- read_excel(
  data_path,
  sheet = "原始数据",
  .name_repair = "minimal"
)

dat <- tibble(
  id = raw[[1]],
  sex = raw[[10]],
  age = raw[[11]],
  
  multi_partner = raw[[12]],
  
  mate1 = raw[[25]],
  mate2 = raw[[26]],
  mate3 = raw[[27]],
  mate4 = raw[[28]],
  mate5 = raw[[29]],
  
  adhd1  = raw[[33]],
  adhd2  = raw[[34]],
  adhd3  = raw[[35]],
  adhd4  = raw[[36]],
  adhd5  = raw[[37]],
  adhd6  = raw[[38]],
  adhd7  = raw[[39]],
  adhd8  = raw[[40]],
  adhd9  = raw[[41]],
  adhd10 = raw[[42]],
  adhd11 = raw[[43]],
  adhd12 = raw[[44]],
  adhd13 = raw[[45]],
  adhd14 = raw[[46]],
  adhd15 = raw[[47]],
  adhd16 = raw[[48]],
  adhd17 = raw[[49]],
  adhd18 = raw[[50]]
)

adhd_map <- c(
  "从不" = 1,
  "很少" = 2,
  "有时" = 3,
  "经常" = 4,
  "非常经常" = 5
)

recode_adhd <- function(x){
  x <- str_squish(as.character(x))
  x_num <- suppressWarnings(as.numeric(x))
  x_lab <- as.numeric(unname(adhd_map[x]))
  ifelse(!is.na(x_num), x_num, x_lab)
}

recode_inter <- function(x){
  x <- str_squish(as.character(x))
  case_when(
    str_detect(x, "^B") ~ 1,
    str_detect(x, "^A") ~ 0,
    TRUE ~ NA_real_
  )
}

row_mean_min <- function(data, vars, min_items = 1){
  x <- data[, vars] %>%
    mutate(across(everything(), as.numeric)) %>%
    as.matrix()
  
  valid_n <- rowSums(!is.na(x))
  out <- rowMeans(x, na.rm = TRUE)
  out[valid_n < min_items] <- NA_real_
  out
}

adhd_items <- paste0("adhd", 1:18)

inatt_items <- c(
  "adhd1", "adhd2", "adhd3", "adhd4",
  "adhd7", "adhd8", "adhd9", "adhd10", "adhd11"
)

hyper_items <- c(
  "adhd5", "adhd6",
  "adhd12", "adhd13", "adhd14", "adhd15",
  "adhd16", "adhd17", "adhd18"
)

inter_items <- paste0("inter", 1:5)
intra_items <- paste0("intra", 1:5)

dat <- dat %>%
  mutate(
    across(all_of(adhd_items), recode_adhd),
    
    inter1 = recode_inter(mate1),
    inter2 = recode_inter(mate2),
    inter3 = recode_inter(mate3),
    inter4 = recode_inter(mate4),
    inter5 = recode_inter(mate5),
    
    intra1 = 1 - inter1,
    intra2 = 1 - inter2,
    intra3 = 1 - inter3,
    intra4 = 1 - inter4,
    intra5 = 1 - inter5
  )

dat <- dat %>%
  mutate(
    adhd_total = row_mean_min(., adhd_items, min_items = 14),
    adhd_inattention = row_mean_min(., inatt_items, min_items = 7),
    adhd_hyperimpulsive = row_mean_min(., hyper_items, min_items = 7),
    
    inter_score = row_mean_min(., inter_items, min_items = 3),
    intra_score = row_mean_min(., intra_items, min_items = 3),
    
    inter_majority = if_else(inter_score >= 0.60, 1, 0, missing = NA_real_),
    intra_majority = if_else(intra_score >= 0.60, 1, 0, missing = NA_real_)
  )

str(dat[, adhd_items])

summary(dat$adhd_total)
summary(dat$inter_score)
summary(dat$intra_score)

table(dat$inter_majority, useNA = "ifany")
table(dat$intra_majority, useNA = "ifany")
```

##CFA model: ADHD as a latent variable

```{r}
dat_cfa <- dat %>%
  select(all_of(adhd_items)) %>%
  drop_na() %>%
  mutate(across(everything(), ordered))

adhd_cfa_total <- paste0(
  "ADHD =~ ",
  paste(adhd_items, collapse = " + ")
)

fit_cfa_total <- cfa(
  adhd_cfa_total,
  data = dat_cfa,
  ordered = adhd_items,
  estimator = "WLSMV",
  parameterization = "theta",
  std.lv = TRUE
)

summary(
  fit_cfa_total,
  fit.measures = TRUE,
  standardized = TRUE
)

fitMeasures(
  fit_cfa_total,
  c("chisq", "df", "pvalue", "cfi", "tli", "rmsea", "srmr")
)

semTools::reliability(fit_cfa_total)
```

Structural model: latent ADHD predicting intersexual-display binary outcome

```{r}
dat_sem_inter <- dat %>%
  select(all_of(adhd_items), inter_majority) %>%
  drop_na() %>%
  mutate(across(everything(), ordered))

sem_inter_model <- paste0(
  "ADHD =~ ",
  paste(adhd_items, collapse = " + "),
  "\ninter_majority ~ ADHD"
)

fit_sem_inter <- sem(
  sem_inter_model,
  data = dat_sem_inter,
  ordered = c(adhd_items, "inter_majority"),
  estimator = "WLSMV",
  parameterization = "theta",
  std.lv = TRUE
)

summary(
  fit_sem_inter,
  fit.measures = TRUE,
  standardized = TRUE
)

parameterEstimates(
  fit_sem_inter,
  standardized = TRUE,
  ci = TRUE
) %>%
  filter(op == "~")
```

Structural model: latent ADHD predicting intrasexual-competition binary outcome

```{r}
dat_sem_intra <- dat %>%
  select(all_of(adhd_items), intra_majority) %>%
  drop_na() %>%
  mutate(across(everything(), ordered))

sem_intra_model <- paste0(
  "ADHD =~ ",
  paste(adhd_items, collapse = " + "),
  "\nintra_majority ~ ADHD"
)

fit_sem_intra <- sem(
  sem_intra_model,
  data = dat_sem_intra,
  ordered = c(adhd_items, "intra_majority"),
  estimator = "WLSMV",
  parameterization = "theta",
  std.lv = TRUE
)

summary(
  fit_sem_intra,
  fit.measures = TRUE,
  standardized = TRUE
)

parameterEstimates(
  fit_sem_intra,
  standardized = TRUE,
  ci = TRUE
) %>%
  filter(op == "~")
```



```{r}
# Function: row mean with minimum number of valid items
row_mean_min <- function(data, cols, min_items = 1) {
  x <- data[, cols, drop = FALSE]
  
  # make sure columns are numeric
  x <- as.data.frame(lapply(x, function(v) as.numeric(as.character(v))))
  
  n_valid <- rowSums(!is.na(x))
  out <- rowMeans(x, na.rm = TRUE)
  
  out[n_valid < min_items] <- NA_real_
  out
}
# Function: row mean with minimum number of valid items
row_mean_min <- function(data, cols, min_items = 1) {
  x <- data[, cols, drop = FALSE]
  
  # make sure columns are numeric
  x <- as.data.frame(lapply(x, function(v) as.numeric(as.character(v))))
  
  n_valid <- rowSums(!is.na(x))
  out <- rowMeans(x, na.rm = TRUE)
  
  out[n_valid < min_items] <- NA_real_
  out
}
```



```{r}
library(dplyr)

# Final analysis dataset: complete-case sample for paper
# This forces the logistic regression, CFA, reliability, and group comparison
# to use the same ADHD item-level complete-case sample.

dat_paper <- dat %>%
  filter(
    if_all(all_of(adhd_items), ~ !is.na(.x)),
    !is.na(inter_majority)
  ) %>%
  mutate(
    adhd_total = rowMeans(across(all_of(adhd_items)), na.rm = FALSE),
    adhd_inattention = rowMeans(across(all_of(inatt_items)), na.rm = FALSE),
    adhd_hyperimpulsive = rowMeans(across(all_of(hyper_items)), na.rm = FALSE)
  )

# Check final sample
n_raw <- nrow(dat)
n_final <- nrow(dat_paper)
n_removed <- n_raw - n_final

n_final

```

```{r}
age_midpoint <- function(x) {
  x <- as.character(x)
  nums <- stringr::str_extract_all(x, "\\d+")
  
  purrr::map_dbl(nums, function(z) {
    z <- as.numeric(z)
    
    if (length(z) >= 2) {
      mean(z[1:2])
    } else if (length(z) == 1) {
      z[1]
    } else {
      NA_real_
    }
  })
}

dat_paper <- dat_paper %>%
  mutate(age_mid = age_midpoint(age))

age_table <- dat_paper %>%
  count(age, name = "n") %>%
  mutate(percent = round(100 * n / sum(n), 2))

knitr::kable(
  age_table,
  caption = "Age category distribution"
)

age_summary <- dat_paper %>%
  summarise(
    n_age = sum(!is.na(age_mid)),
    mean_age = mean(age_mid, na.rm = TRUE),
    sd_age = sd(age_mid, na.rm = TRUE),
    min_age = min(age_mid, na.rm = TRUE),
    max_age = max(age_mid, na.rm = TRUE)
  ) %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))

knitr::kable(
  age_summary,
  caption = "Approximate age summary using category midpoints"
)
```

##ADHD reliability: alpha and omega
```{r}
#| message: false
#| warning: false

# Reset item names to match your current dataset
adhd_items <- paste0("adhd", 1:18)

inatt_items <- c(
  "adhd1", "adhd2", "adhd3", "adhd4",
  "adhd7", "adhd8", "adhd9", "adhd10", "adhd11"
)

hyper_items <- c(
  "adhd5", "adhd6",
  "adhd12", "adhd13", "adhd14", "adhd15",
  "adhd16", "adhd17", "adhd18"
)

# Check that the columns exist
stopifnot(all(adhd_items %in% names(dat_paper)))
stopifnot(all(inatt_items %in% names(dat_paper)))
stopifnot(all(hyper_items %in% names(dat_paper)))

# Reliability dataset
adhd_reliability_dat <- dat_paper %>%
  select(all_of(adhd_items)) %>%
  drop_na()

# Cronbach's alpha
alpha_total <- psych::alpha(adhd_reliability_dat)
alpha_inattention <- psych::alpha(
  adhd_reliability_dat %>% select(all_of(inatt_items))
)
alpha_hyper <- psych::alpha(
  adhd_reliability_dat %>% select(all_of(hyper_items))
)

# Omega
omega_total <- psych::omega(
  adhd_reliability_dat,
  nfactors = 2,
  plot = FALSE
)

# Clean reliability table
reliability_table <- tibble(
  Scale = c(
    "ADHD total",
    "Inattention",
    "Hyperactivity / impulsivity",
    "ADHD total omega total",
    "ADHD total omega hierarchical"
  ),
  Estimate = c(
    alpha_total$total$raw_alpha,
    alpha_inattention$total$raw_alpha,
    alpha_hyper$total$raw_alpha,
    omega_total$omega.tot,
    omega_total$omega_h
  )
) %>%
  mutate(Estimate = round(Estimate, 3))

knitr::kable(
  reliability_table,
  caption = "Reliability estimates for ADHD items"
)
```
## two factor ADHD CFA
```{r}
dat_cfa_2factor <- dat_paper %>%
  select(all_of(adhd_items)) %>%
  drop_na() %>%
  mutate(across(everything(), ordered))

adhd_2factor_model <- '
  Inattention =~ adhd1 + adhd2 + adhd3 + adhd4 +
                 adhd7 + adhd8 + adhd9 + adhd10 + adhd11

  HyperImp =~ adhd5 + adhd6 + adhd12 + adhd13 + adhd14 +
              adhd15 + adhd16 + adhd17 + adhd18

  Inattention ~~ HyperImp
'

fit_adhd_2factor <- cfa(
  model = adhd_2factor_model,
  data = dat_cfa_2factor,
  ordered = adhd_items,
  estimator = "WLSMV",
  parameterization = "theta",
  std.lv = TRUE
)

fit_2factor_table <- tibble(
  Model = "Two-factor ADHD",
  chisq = fitMeasures(fit_adhd_2factor, "chisq"),
  df = fitMeasures(fit_adhd_2factor, "df"),
  CFI = fitMeasures(fit_adhd_2factor, "cfi"),
  TLI = fitMeasures(fit_adhd_2factor, "tli"),
  RMSEA = fitMeasures(fit_adhd_2factor, "rmsea"),
  SRMR = fitMeasures(fit_adhd_2factor, "srmr")
) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3)))

knitr::kable(
  fit_2factor_table,
  caption = "Two-factor ADHD CFA model fit"
)
```
## comparison
```{r}
knitr::kable(
  fit_2factor_table,
  caption = "Two-factor ADHD CFA model fit"
)

adhd_2factor_model <- '
  Inattention =~ adhd1 + adhd2 + adhd3 + adhd4 +
                 adhd7 + adhd8 + adhd9 + adhd10 + adhd11

  HyperImp =~ adhd5 + adhd6 + adhd12 + adhd13 + adhd14 +
              adhd15 + adhd16 + adhd17 + adhd18

  Inattention ~~ HyperImp
'
knitr::kable(
  fit_2factor_table,
  caption = "Two-factor ADHD CFA model fit"
)
## Figure 1. Two-factor CFA path diagram

#| label: fig-cfa-path
#| fig.cap: "Two-factor CFA measurement model for ADHD symptoms. Inattention and Hyperactivity/Impulsivity were modeled as correlated latent factors. Each factor loaded onto its corresponding ASRS-based observed items."
#| fig.width: 9
#| fig.height: 5
#| fig.align: center
#| message: false
#| warning: false

library(ggplot2)
library(dplyr)
library(grid)

# Observed item positions
items <- tibble(
  factor = rep(c("Inattention", "Hyperactivity / Impulsivity"), each = 9),
  item = c(
    "adhd1", "adhd2", "adhd3", "adhd4", "adhd7", "adhd8", "adhd9", "adhd10", "adhd11",
    "adhd5", "adhd6", "adhd12", "adhd13", "adhd14", "adhd15", "adhd16", "adhd17", "adhd18"
  ),
  x = rep(seq(-4, 4, length.out = 9), 2),
  y = c(rep(1.1, 9), rep(-1.1, 9)),
  latent_x = 0,
  latent_y = c(rep(2.0, 9), rep(0.0, 9))
)

# Latent factor labels
latent <- tibble(
  name = c("Inattention", "Hyperactivity /\nImpulsivity"),
  x = c(0, 0),
  y = c(2.2, 0)
)

ggplot() +
  # Factor loading arrows
  geom_segment(
    data = items,
    aes(x = latent_x, y = latent_y, xend = x, yend = y + 0.18),
    arrow = arrow(length = unit(0.12, "inches")),
    linewidth = 0.35
  ) +
  
  # Correlation between latent factors: two curved arrows
  geom_curve(
    aes(x = 0.55, y = 1.85, xend = 0.55, yend = 0.35),
    curvature = -0.25,
    arrow = arrow(length = unit(0.12, "inches")),
    linewidth = 0.35
  ) +
  geom_curve(
    aes(x = -0.55, y = 0.35, xend = -0.55, yend = 1.85),
    curvature = -0.25,
    arrow = arrow(length = unit(0.12, "inches")),
    linewidth = 0.35
  ) +
  
  # Observed item boxes
  geom_label(
    data = items,
    aes(x = x, y = y, label = item),
    fill = "white",
    label.size = 0.30,
    size = 3.2
  ) +
  
  # Latent variable circles
  annotate("point", x = 0, y = 2.2, shape = 21, size = 32, fill = "white", stroke = 0.8) +
  annotate("text", x = 0, y = 2.2, label = "Inattention", size = 3.4) +
  annotate("point", x = 0, y = 0, shape = 21, size = 32, fill = "white", stroke = 0.8) +
  annotate("text", x = 0, y = 0, label = "Hyperactivity /\nImpulsivity", size = 3.1) +
  
  coord_cartesian(xlim = c(-4.8, 4.8), ylim = c(-1.6, 2.7), clip = "off") +
  theme_void()
```
```
## logistic regression
```{r}
dat_logit <- dat_paper %>%
  mutate(
    adhd_z = as.numeric(scale(adhd_total))
  ) %>%
  select(adhd_z, inter_majority) %>%
  drop_na()

m_inter_total <- glm(
  inter_majority ~ adhd_z,
  data = dat_logit,
  family = binomial(link = "logit")
)

logit_table <- broom::tidy(
  m_inter_total,
  conf.int = TRUE,
  exponentiate = FALSE
) %>%
  mutate(
    OR = exp(estimate),
    OR_low = exp(conf.low),
    OR_high = exp(conf.high)
  ) %>%
  select(
    Predictor = term,
    B = estimate,
    SE = std.error,
    z = statistic,
    p = p.value,
    OR,
    OR_low,
    OR_high
  ) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3)))

knitr::kable(
  logit_table,
  caption = "Logistic regression predicting intersexual / ornament-like majority choice"
)
```

## diagnostics table
```{r}
m_null <- glm(
  inter_majority ~ 1,
  data = dat_logit,
  family = binomial(link = "logit")
)

lrt_table <- anova(m_null, m_inter_total, test = "Chisq")

knitr::kable(
  lrt_table,
  caption = "Likelihood-ratio test comparing null model and ADHD_total model"
)

diagnostic_table <- tibble(
  Diagnostic = c(
    "Model convergence",
    "Outcome type",
    "Estimator",
    "Multicollinearity"
  ),
  Result = c(
    ifelse(m_inter_total$converged, "Converged", "Did not converge"),
    "Binary outcome: inter_majority coded 0/1",
    "Maximum likelihood estimation via glm() with binomial logit link",
    "Not a concern because the focal model has one predictor"
  )
)

knitr::kable(
  diagnostic_table,
  caption = "Model diagnostics and analytical choices"
)
```
## group 
```{r}
#| message: false
#| warning: false
#| fig.width: 6
#| fig.height: 4

# Make sure ggplot2 is available
library(ggplot2)

# Create ADHD median-split group
dat_group <- dat_paper %>%
  mutate(
    inter_majority_num = as.numeric(as.character(inter_majority)),
    adhd_group = if_else(
      adhd_total >= median(adhd_total, na.rm = TRUE),
      "Higher ADHD group",
      "Lower ADHD group"
    )
  ) %>%
  filter(!is.na(inter_majority_num), !is.na(adhd_group))

# Group summary with 95% CI
group_summary <- dat_group %>%
  group_by(adhd_group) %>%
  summarise(
    n = n(),
    proportion = mean(inter_majority_num, na.rm = TRUE),
    se = sqrt(proportion * (1 - proportion) / n),
    lower = pmax(0, proportion - 1.96 * se),
    upper = pmin(1, proportion + 1.96 * se),
    .groups = "drop"
  )

# Table
knitr::kable(
  group_summary %>%
    mutate(
      proportion = round(proportion, 3),
      se = round(se, 3),
      lower = round(lower, 3),
      upper = round(upper, 3)
    ),
  caption = "Descriptive group comparison with 95% confidence intervals"
)

# Plot
ggplot(group_summary, aes(x = adhd_group, y = proportion)) +
  geom_col(width = 0.60) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.15) +
  scale_y_continuous(
    limits = c(0, 1),
    labels = function(x) paste0(round(x * 100), "%")
  ) +
  labs(
    x = NULL,
    y = "Intersexual / ornament-like majority choice",
    title = "Intersexual choice by ADHD group"
  ) +
  theme_minimal()
```






